Array

### 2014-03-02 DOCUMENTATION/AdvancedTypes.tex
Adding structure support details.
The next step is review the AdvancedTypes documentation

	\multicolumn{3}{c}{ \B{Vari} \alt \B{ArrayItem} }\\

@@

        \multicolumn{3}{c} {\B{Lit} \alt \B{Term} \alt \B{Pred} \alt \B{Cons}}\\
	\multicolumn{3}{c} {\B{Vari} \alt \B{Array} \alt \B{ArrayItem} } \\

@@

@@

\begin{tabular}[t]{rcl}
        \hline
        \multicolumn{3}{|c|}{Comparison: \B{Comp}} \\
        \hline
        \B{op} & : & \B{RelOp} \\
        \B{arg1}, \B{arg2} & : & \B{Exp} \\
        \hline
        \multicolumn{3}{|c|}{Data type: \B{Type}} \\
        \hline
        \multicolumn{3}{c}{\B{INT} \alt \B{BOOL}\alt}\\
	\multicolumn{3}{c}{\B{Struct} \alt \B{ArrayType}}\\
	\hline
  	\multicolumn{3}{|c|}{Data structure : \B{Struct}} \\
        \hline
	\B{id} & : &  \B{Name}\\
        \B{elements} & : & seq \B{Name} \\
	\B{type} & : & seq \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Access struct : \B{StructItem}} \\
        \hline
	\B{base} & : &  \B{Struct} \\
	\B{element} & : & \B{Name}\\
	\B{type} & : & \B{Type}\\
	\hline
        \multicolumn{3}{|c|}{Array type : \B{ArrayType}} \\
        \hline 
	\B{dom} & : & seq \B{Set}\\
	\B{ran} & : & \B{Set}\\
        \hline
        \multicolumn{3}{|c|}{Array expression : \B{Array}} \\
        \hline
	\multicolumn{3}{c}{ seq \B{Map} \alt \B{Product} \alt \B{Name}  }\\
	\hline
  	\multicolumn{3}{|c|}{Ordered pair : \B{Map}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 
	\hline
  	\multicolumn{3}{|c|}{Cartesian product : \B{Product}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 	
	\hline
	\multicolumn{3}{|c|}{Access array : \B{ArrayItem}} \\
        \hline
	\B{base} & : &  \B{Array} \\
	\B{indexes} & : & seq \B{Expr}\\
	\B{type} & : & \B{ArrayType}\\
	\hline
	\multicolumn{3}{|c|}{Simple set : \B{Set}} \\
	\hline
	\multicolumn{3}{c}{\B{BOOL} \alt \B{NAT} \alt \B{NAT$_{1}$}\alt } \\
	\multicolumn{3}{c}{\B{INT} \alt \B{Interval}} \\
	\hline
	\multicolumn{3}{|c|}{B0 interval : \B{Interval}} \\
	\hline
	\B{start}, \B{end} & : & \B{Expr} \\
	\hline
	\multicolumn{3}{|c|}{Literal : \B{Lit}} \\
	\hline
	\multicolumn{3}{c}{ \B{Bool\_lit} \alt \B{Int\_lit}} \\ 
      \end{tabular}
    \end{tabular}
  }
  }
    \caption{Abstract syntax structure of a B implementation.  For each abstract
      syntax element, we give the name of the class (e.g. \B{CaseBr} for a
      branch in a case instruction, and the different attributes of the class),
      or a list of the possible sub-classes (e.g. \B{Inst} for the different
      kinds of instructions).
      Sequence and optional attributes are denoted ``seq'' and ``opt'',
      respectively}
    \label{tab:node-attr}
  \end{center}
\end{figure}

@@

\section{Translation of Array and Structures}

@@

\section{Structure}
The struct declaration defines a list of variables to be placed under one name, 
allowing several variables to be accessed by one base name. This example get andset one value of xx'balance and initializes a structure.

@@

The following rule get an element value from the struct. The same rule is used in left side of attribution and right side of attribution. The functions \SBPos(n) and \SBType(n)  get respectively the position and type of $n$ element in the structure. 
\begin{align*}
\begin{split}
  \trad{n}{\B{StructItem}} \isdef \\
   & \LET \ell = \lbl \AND \PH{t} = n.type \AND \PH{v} = n.element \IN\\
   & \llvm{"\PH{\ell} = getelementptr \%\PH{t} @\PH{v}, i32 0, i32  \SBPos(\B{n})"} \\
   & \sep  \ell  \sep \SBType(\B{n})  \\
\end{split}
\end{align*}

AST

record structure : Record
elements : seq Name;
values : seq Expr;

Isto é uma expressão

To remove the restrictions:
( ) The direct attribution between arrays  is madded in similar way

\section{Arrays}
%TODO: update according the new names in AST, just change the names.
A concrete array is a total function where the domain indexes the range 
elements. The domain is composed of Cartesian product of sequence of concrete 
sets. The node Array in AST represents a declaration of array or a 
selection of element. The first element ``dec'' declares explicitly an array or 
references a variable array. The second element ``type'' defines a sequence of 
array types, where the last element defines the range of function and the other 
elements define the domain. The third element ``indexes'' is optional, and it 
represents the selection of an element in array, or the same type structure 
when it is a structure.

@@

\subsection{Rules}

The rule lvalue is:
\begin{align*}
\lefteqn{\trad{n}{lvalue} \isdef} \\
& \LET t = \llvm{"\PH{\trad{\B{n.type}}{Type}}"} \IN \\
& \quad \IF \mbox{\B{n} is a local variable} \THEN \llvm{""} \sep \local(\B{n}) \sep t \\
& \quad \IF \mbox{\B{n} is an array} \THEN  \trad{n}{ArrayItem} \\
& \quad \ELSE \LET v = \variable \IN \\
& \quad \quad \llvm{"\PH{v} = getelementptr \PH{\state(\B{n.root})}* \%self\$,} \\
& \quad \quad \quad \llvm{i32 0, i32 \PH{\idx(\B{n})} \nl"} \sep v \sep t
\end{align*}

@@

\paragraph{Restriction on array:} The translation of direct attribution
 between array expression is not yet supported, for example:
\begin{pascalcode}
 array1 := array2;
\end{pascalcode}

>>>>

	\multicolumn{3}{c}{ \B{Vari} \alt \B{ArrayItem} \alt \B{StructItem} }\\

@@

        \multicolumn{3}{c}{\B{Lit}\alt\B{Term}\alt\B{Pred}\alt\B{Cons}\alt\B{Record}}\\
	\multicolumn{3}{c}{\B{Vari}\alt\B{StructItem}\alt\B{ArrayItem} } \\

@@

        \multicolumn{3}{|c|}{Comparison: \B{Comp}} \\
        \B{op} & : & \B{RelOp} \\
        \B{arg1}, \B{arg2} & : & \B{Exp} \\

@@

\begin{tabular}[t]{rcl}
        \hline
        \multicolumn{3}{|c|}{Data type: \B{Type}} \\
        \hline
        \multicolumn{3}{c}{\B{INT} \alt \B{BOOL}\alt}\\
	\multicolumn{3}{c}{\B{Struct} \alt \B{ArrayType}}\\
	\hline
  	\multicolumn{3}{|c|}{Data structure : \B{Struct}} \\
        \hline
	\B{id} & : &  \B{Name}\\
        \B{elements} & : & seq \B{Name} \\
	\B{type} & : & seq \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Access struct : \B{StructItem}} \\
        \hline
	\B{base} & : &  \B{Struct} \\
	\B{element} & : & \B{Name}\\
	\B{type} & : & \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Record Structure : \B{Record}} \\
        \hline
	\B{elements} & : & seq \B{Name}\\
	\B{values} & : & seq \B{Expr}\\
	\hline
        \multicolumn{3}{|c|}{Array type : \B{ArrayType}} \\
        \hline 
	\B{dom} & : & seq \B{Set}\\
	\B{ran} & : & \B{Set}\\
        \hline
        \multicolumn{3}{|c|}{Array expression : \B{Array}} \\
        \hline
	\multicolumn{3}{c}{ seq \B{Map} \alt \B{Product} \alt \B{Name}  }\\
	\hline
  	\multicolumn{3}{|c|}{Ordered pair : \B{Map}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 
	\hline
  	\multicolumn{3}{|c|}{Cartesian product : \B{Product}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 	
	\hline
	\multicolumn{3}{|c|}{Access array : \B{ArrayItem}} \\
        \hline
	\B{base} & : &  \B{Array} \\
	\B{indexes} & : & seq \B{Expr}\\
	\B{type} & : & \B{ArrayType}\\
	\hline
	\multicolumn{3}{|c|}{Simple set : \B{Set}} \\
	\hline
	\multicolumn{3}{c}{\B{BOOL} \alt \B{NAT} \alt \B{NAT$_{1}$}\alt } \\
	\multicolumn{3}{c}{\B{INT} \alt \B{Interval}} \\
	\hline
	\multicolumn{3}{|c|}{B0 interval : \B{Interval}} \\
	\hline
	\B{start}, \B{end} & : & \B{Expr} \\
	\hline
	\multicolumn{3}{|c|}{Literal : \B{Lit}} \\
	\hline
	\multicolumn{3}{c}{ \B{Bool\_lit} \alt \B{Int\_lit}} \\ 
      \end{tabular}
    \end{tabular}
  }
  }
    \caption{Abstract syntax structure of a B implementation.  For each abstract
      syntax element, we give the name of the class (e.g. \B{CaseBr} for a
      branch in a case instruction, and the different attributes of the class),
      or a list of the possible sub-classes (e.g. \B{Inst} for the different
      kinds of instructions).
      Sequence and optional attributes are denoted ``seq'' and ``opt'',
      respectively}
    \label{tab:node-attr}
  \end{center}
\end{figure}

@@

\subsection{Rules to be updated in first specification}

The rule lvalue is:
\begin{align*}
\lefteqn{\trad{n}{lvalue} \isdef} \\
& \LET t = \llvm{"\PH{\trad{\B{n.type}}{Type}}"} \IN \\
& \quad \IF \mbox{\B{n} is a local variable} \THEN \llvm{""} \sep \local(\B{n}) \sep t \\
& \quad \IF \mbox{\B{n} is an array item} \THEN  \trad{n}{ArrayItem} \\
& \quad \IF \mbox{\B{n} is a struct item} \THEN  \trad{n}{StructItem} \\
& \quad \ELSE \LET v = \variable \IN \\
& \quad \quad \llvm{"\PH{v} = getelementptr \PH{\state(\B{n.root})}* \%self\$,} \\
& \quad \quad \quad \llvm{i32 0, i32 \PH{\idx(\B{n})} \nl"} \sep v \sep t
\end{align*}

\section{Translation of Structures and Array}

@@

\subsection{Structure}
The struct declaration defines a list of variables to be placed under one name, 
allowing several variables to be accessed by one base name. This example get andset one value of xx'balance and initializes a structure.

@@

The following rule get an element value from the struct. The same rule is used in left side of attribution as a $LValue$ and right side of attribution as an $Expression$. The functions \SBPos(n) and \SBType(n)  get respectively the position and type of $n$ element in the structure.

\begin{align*}
\begin{split}
  \trad{n}{\B{StructItem}} \isdef \\
   & \LET \ell = \lbl \AND \PH{t} = n.type \AND \PH{v} = n.element \IN\\
   & \llvm{"\PH{\ell} = getelementptr \%\PH{t} @\PH{v}, i32 0, i32  \SBPos(\B{n})"} \\
   & \sep  \ell  \sep \SBType(\B{n})  \\
\end{split}
\end{align*}

The attribution from record structure to a structure variable, line 1 from last code example, is translated by a sequence of application of rule $StructItem$.

\section{Arrays}
A concrete array is a total function where the domain indexes the range 
elements. The domain is composed of Cartesian product of sequence of concrete 
sets. The node Array in AST represents a declaration of array or a 
selection of element.

@@

\subsection{Rules for support Array}

@@

\paragraph{Restriction on array:} The translation of direct attribution
 between array expression is not yet supported, for example:

\begin{pascalcode}
 array1 := array2;
\end{pascalcode}
